---
date: 2017-04-19T06:25:41+02:00
title: "Scraping data from STATEC's public tables"
author: Bruno Rodrigues
tags: R
---

A lot of open data is available in Luxembourg's [open data portal](https://data.public.lu/en/), but
sometimes, it is not very easy to download. In the video below, I give you an example of such data and
show how you can use `rvest` to get the data easily.

After watching the video, take a look at the code below. This code does two things; first it scraps
the data as I show in the video, and then it puts the data in a tidy format fur further processing.



So to summarize the idea of the video; instead of clicking the buttons to download each year's data
(which you would have to do 15 times), it is easier to simple turn off javascript and then scrap the
html version of the table. It would be possible, albeit with much more effort, to scrap the tables
with javascript enabled, by using a tool such as [phantomjs](http://phantomjs.org/). But since we have
the possibility to view the table in html, why not take advantage of it?


To scrap the data, you will need first to install the `rvest` and then load it (and let's also load
the other needed packages)

```{r}
library(rvest)
library(dplyr)
library(purrr)
```

```{r, cache = TRUE}
page_unemp <- read_html("http://www.statistiques.public.lu/stat/TableViewer/tableViewHTML.aspx?ReportId=12950&IF_Language=eng&MainTheme=2&FldrName=3&RFPath=91")

page_unemp %>%
  html_nodes(".b2020-datatable") %>% .[[1]] %>% html_table(fill = TRUE) -> data_unemp


head(data_unemp)
```

As you can see, we got the data in quite a nice format, but it still needs to be cleaned a bit. Let's
do this.

First, let's use the first row as the header of the data set and then remove it:
```{r}
colnames(data_unemp) <- data_unemp[2, ]
colnames(data_unemp)[1:2] <- c("division", "variable")
data_unemp <- data_unemp[-c(1,2), ]
head(data_unemp)
```
This is starting to look nice, but we need to replace the "," with "." and then convert the columns
to numeric.

```{r}
data_unemp %>%
  map_df(function(x)(gsub(",", ".", x = x))) %>%
  mutate_at(vars(matches("\\d{4}")), as.numeric
            ) -> clean_unemp

head(clean_unemp)
```

This line `map_df(function(x)(gsub(",", ".", x = x)))` calls `purrr::map_df()`, which maps a function
to each column of a data frame. The function in question is `function(x)(gsub(",", ".", x = x))`,
which is an anonymous function (meaning it does not have a name). This function looks for the string
"," and replaces it with "." in a single column of the data frame. But because we're mapping this
function to all the columns of the data frame with `purrr::map_df()`, this substitution happens
in each column. We' not done yet, because these columns are still holding characters. We need to
convert each column to a numeric vector and this is what happens in the next line, `mutate_at(vars(matches("200*")), as.numeric)`.
Each column that starts with the string "20" is converted to numeric with `dplyr::mutate_at()`.
